---
title: "Using c3co after PSCBS Segmentation"
author: "Morgane Pierre-Jean and Pierre Neuvial"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: c3co.bib
nocite: | 
  @olshen11parent-specific

vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using c3co after PSCBS Segmentation}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This document provides a brief tutorial for using the c3co package on data segmented by PSCBS (@olshen11parent-specific).

```{r}
library("c3co")
library("c3co.data")
```

## Using wrapper

In this document we run c3co on a public data set in the [PSCBS] format, which has been saved into the [c3co.data] package. The `PSCBSwrapper()` function segments and saves data from PSCBS format to c3co format.

```{r, eval=TRUE}
data("PSCBSdata", package = "c3co.data")
segDat <- PSCBSwrapper(PSCBSdata, stat = "C1C2")
```

## Run c3co ##
Now, you can run c3co by using this line:
```{r, message=FALSE, eval=TRUE, warning=FALSE}
lambda.grid <- seq(from = 1e-4, to = 1e-3, length.out = 10L)
parameters.grid <- list(lambda = lambda.grid, nb.arch = 2:5)
resc3co <- c3co(dat = NULL, parameters.grid, segDat = segDat)
```
```{r, fig.width=7, fig.height=5, eval=TRUE}
pvePlot(resc3co)
```

```{r, fig.width=7, fig.height=5, eval=TRUE}
Wplot(resc3co, idxBest = 3L)
```

```{r, fig.width=7, fig.height=5, eval=TRUE}
df <- createZdf(resc3co, chromosomes = 1:2, idxBest = 2)
```

```{r, fig.width=7, fig.height=5, eval=TRUE}
Zplot(df, scalePosToMb = TRUE)
```

## Session Information

```{r}
sessionInfo()
```


## References


[c3co.data]: https://github.com/pneuvial/c3co.data
[PSCBS]: https://cran.r-project.org/package=PSCBS
